---
title: "Project goal and data description"
author: "Eloïse Vannier"
---

```{r}
#| message: false
here::i_am("final_project.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
```

You can find the github repository associated to this project via this link : [github-repository](https://github.com/E10is3/final-project.git)

## Sources :

The data I use come from two open source data sets of the World Bank. The first one includes governance indicators available via this link : [WorldBank-governance-indicators](https://databank.worldbank.org/source/worldwide-governance-indicators#)

Please note that the ticked options are the preset ones by the website and not the ones I will use. To get the exact same .csv file, select all the 214 countries available, the years 2020, 2021 and 2022, and all the 36 variable series before the download.

The second data set includes development indicators, available via this link : [WorldBank-development-indicators](https://databank.worldbank.org/source/world-development-indicators#)

To get the exact same .csv file, select all the 266 countries available, the years 2020, 2021 and 2022, and the 29 variable series listed in the description below.

## Description of the sources :

#### Governance indicators :

The data set include indicators for six dimensions of governance: Control of Corruption; Government Effectiveness; Political Stability and Absence of Violence/Terrorism; Regulatory Quality; Rule of Law; and Voice and Accountability. For each dimensions, the data set provides an indicator and its standard error, the rank of the country, both upper and lower bound of a 95% confidence interval for this ranking, as well as the number of sources used to compute the indicator.
#### Development indicators : 

The data set is a compilation the World Bank did from officially recognized international sources, including national, regional and global estimates. In order to have a multidimensional perception of the development of each country, I selected well known indicators (GDP per capita, GDP PPP, FDI inflows and outflows) as well as indicators related to health such as hospital bed per 1000 people or life expectancy; education such as literacy rate or children out of school; infrastructure such as access to electricity or internet use and poverty.

## Data loading :

```{r}
#| message: false
govc_indicators <- vroom(here("data", "governance_indicators.csv"))
dvpt_indicators <- vroom(here("data", "development_indicators.csv"))
```

```{r}
govc_indicators |> summarise("nb of lines" = NROW(govc_indicators),"nb of colmuns"= NCOL(govc_indicators))|>
  knitr::kable(align = 'c', caption = "Governance data set properties")
```

```{r}
dvpt_indicators |> summarise("nb of lines" = NROW(dvpt_indicators),"nb of colmuns"= NCOL(dvpt_indicators))|>
  knitr::kable(align = 'c', caption = "Development data set properties")
```

## Data cleaning and joining :

### Dealing with the missing variable :
In both data set, a lot of cells contain "..", that can easily be interpreted as missing value. We need to take this into account :

```{r}
dvpt_indicators [dvpt_indicators == ".."] <- NA
govc_indicators [govc_indicators == ".."] <- NA
```



