---
title: "Project goal and data description"
author: "Eloïse Vannier"
---

```{r}
#| message: false
here::i_am("final_project.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
```

You can find the github repository associated to this project via this link : [github-repository](https://github.com/E10is3/final-project.git)

## Sources :

The data I use come from two open source data sets of the World Bank. The first one includes governance indicators available via this link : [WorldBank-governance-indicators](https://databank.worldbank.org/source/worldwide-governance-indicators#)

Please note that the ticked options are the preset ones by the website and not the ones I will use. To get the exact same .csv file, select all the 214 countries available, the years 2020, 2021 and 2022, and all the 36 variable series before the download.

The second data set includes development indicators, available via this link : [WorldBank-development-indicators](https://databank.worldbank.org/source/world-development-indicators#)

To get the exact same .csv file, select all the 266 countries available, the years 2020, 2021 and 2022, and the 29 variable series listed in the description below.

## Description of the sources :

#### Governance indicators :

The data set include indicators for six dimensions of governance: Control of Corruption; Government Effectiveness; Political Stability and Absence of Violence/Terrorism; Regulatory Quality; Rule of Law; and Voice and Accountability. For each dimensions, the data set provides an indicator and its standard error, the rank of the country, both upper and lower bound of a 95% confidence interval for this ranking, as well as the number of sources used to compute the indicator.

#### Development indicators : 

The data set is a compilation the World Bank did from officially recognized international sources, including national, regional and global estimates. In order to have a multidimensional perception of the development of each country, I selected well known indicators (GDP per capita, GDP PPP, FDI inflows and outflows) as well as indicators related to health such as hospital bed per 1000 people or life expectancy; education such as literacy rate or children out of school; infrastructure such as access to electricity or internet use and poverty.

## Data loading :

```{r}
govc_indicators <- vroom(here("data", "governance_indicators.csv"))
dvpt_indicators <- vroom(here("data", "development_indicators.csv"))
```
We can see that the columns containing the data of 2020, 2021 and 2022 are not all of the type double but character, we will need to tackle this issue later to be able to study the data sets.

```{r}
#| message: false
govc_indicators |> summarise("nb of lines" = NROW(govc_indicators),"nb of colmuns"= NCOL(govc_indicators))|>
  knitr::kable(align = 'c', caption = "Governance data set properties")
```

```{r}
#| message: false
dvpt_indicators |> summarise("nb of lines" = NROW(dvpt_indicators),"nb of colmuns"= NCOL(dvpt_indicators))|>
  knitr::kable(align = 'c', caption = "Development data set properties")
```

## Data cleaning and joining :

### Dealing with the missing variable :
In both data set, a lot of cells contain "..", that can easily be interpreted as missing value. We need to take this into account :

```{r}
#| message: false
dvpt_indicators [dvpt_indicators == ".."] <- NA
govc_indicators [govc_indicators == ".."] <- NA
```

### Keeping only rows with countries data :

The world bank provided information about data collection and definition at the end of each data set. We know that the country code of the rows that contain countries data are 3 characters so we can select them for the data set with the smallest number of countries/rows :
```{r}
#| message: false
govc_indicators = govc_indicators |> filter(nchar(`Country Code`)==3)
```

Then we can keep the list of the countries and their code in a separate data frame.
```{r}
#| message: false
list_country = govc_indicators |> distinct (`Country Name`,`Country Code`)
```

And select in the development data frame the data of the corresponding countries.
```{r}
#| message: false
dvpt_indicators = inner_join(list_country,dvpt_indicators,join_by(`Country Name`,`Country Code`))
```

### Changing strings into numbers

We saw just after loading the data that type of the columns containing the data of 2020, 2021 and 2022 was character (except for the column 2022 of the governance data frame). We make sure that we read those data as numerical values and not as strings :
```{r}
#| message: false
govc_indicators <- govc_indicators |>  mutate(`2021 [YR2021]` = parse_double(`2021 [YR2021]`),`2020 [YR2020]` = parse_double(`2020 [YR2020]`))

dvpt_indicators <- dvpt_indicators |>  mutate(`2022 [YR2022]` = parse_double(`2022 [YR2022]`), `2021 [YR2021]` = parse_double(`2021 [YR2021]`),`2020 [YR2020]` = parse_double(`2020 [YR2020]`))
```

### Combining the 3 years columns into one :

For countries, indicators values are only available for one or two years. 
We compute the mean of those three columns to have a variable column with way less NA.
```{r}
#| message: false
mean.govc <- rowMeans((govc_indicators |> select(`2020 [YR2020]`,`2021 [YR2021]`, `2022 [YR2022]`)),na.rm = TRUE)
govc_indicators <- govc_indicators |> mutate("Mean"=mean.govc)
mean.dvpt <- rowMeans((dvpt_indicators |> select(`2020 [YR2020]`,`2021 [YR2021]`, `2022 [YR2022]`)),na.rm = TRUE)
dvpt_indicators <- dvpt_indicators |> mutate("Mean"=mean.dvpt)
```

### Saving the variables code in a separate data frame for both data sets:

The "series codes" of the data sets don't seem useful at the moment so we save them in another data frame. 
```{r}
#| message: false
govc_codes = govc_indicators |> distinct(`Series Name`,`Series Code`) 
dvpt_codes = dvpt_indicators |> distinct(`Series Name`,`Series Code`)
```

We can now suppress this column in both data frame, as well as the three year columns. 
```{r}
#| message: false
govc_indicators = govc_indicators |> select(-`Series Code`,-`2020 [YR2020]`,-`2021 [YR2021]`, -`2022 [YR2022]`)
dvpt_indicators = dvpt_indicators |> select(-`Series Code`,-`2020 [YR2020]`,-`2021 [YR2021]`, -`2022 [YR2022]`)
```

### Removing the indicators for which there are too many NA :

Let's arbitrary say that an indicator is only interesting if it is measured for at least 75% of the sample of countries. We first compute this value :
```{r}
#| message: false
critic_minimum = 0.75*NROW(list_country)
```

We then do a list of indicators that meet the criteria :
```{r}
#| message: false
keep.dvpt.indicator = " "
for (i in 1:NROW(dvpt_codes)) {
  df = dvpt_indicators |> filter(dvpt_indicators$`Series Name`== dvpt_codes$`Series Name`[i])
  if (sum(is.na(df$Mean)) < critic_minimum) {
    keep.dvpt.indicator = c(keep.dvpt.indicator,dvpt_codes$`Series Name`[i])
  }
}
```

And we modify the data set accordingly :
```{r}
#| message: false
keep.dvpt.indicator = as.data.frame(keep.dvpt.indicator) |>
  rename(`Series Name`=keep.dvpt.indicator)
dvpt_indicators <- inner_join(keep.dvpt.indicator,dvpt_indicators,join_by(`Series Name`))
```
We kept `r NROW(keep.dvpt.indicator)-1` indicators out of the `r NROW(dvpt_codes)` we selected in the first place.

We do the same for the governance data set :
```{r}
#| message: false
keep.govc.indicator = " "
for (i in 1:NROW(govc_codes)) {
  df = govc_indicators |> filter(govc_indicators$`Series Name`== govc_codes$`Series Name`[i])
  if (sum(is.na(df$Mean)) < critic_minimum) {
    keep.govc.indicator = c(keep.govc.indicator,govc_codes$`Series Name`[i])
  }
}
keep.govc.indicator = as.data.frame(keep.govc.indicator) |>
  rename(`Series Name`=keep.govc.indicator)
govc_indicators <- inner_join(keep.govc.indicator,govc_indicators,join_by(`Series Name`))
```
We kept `r NROW(keep.govc.indicator)-1` indicators and there were `r NROW(govc_codes)`selected in the first place : all of them were well documented across the sample of countries.

## Research question :
The goal of this project is to study the links between the governance of a country, and its development. 
The notion of development is interesting as it can approximate well-being.It encompasses economic aspects, as poverty has a negative impact on the happiness for sure, but can't be limited to this economic lens. More and more economists, sociologist and philosophers highlight the limits of the GDP as indicator of well-being. Indeed, economic development alone may leads to disparities, leaving certain segments of the population marginalized. It is therefore crucial to take into account other dimensions of human development such as access to a good health system and education, or social inequalities. 
Since the robustness of legal frameworks and the protection of human rights can also impact well-being (increasing a feeling of safety and justice), it is logical to add governance aspects to this study.
Governance refers to the systems, structures, and processes by which a country is directed and controlled. It encompasses not only formal government institutions but also all practices that influence decision-making. Effective governance is a cornerstone of development, as it establishes the rules of the game, ensures accountability, and promotes transparency. The quality of governance profoundly influences a country's attractiveness to investors, the efficiency of public services, and the overall well-being of its citizens.
Therefore, understanding these links is interesting for designing effective policies, and fostering sustainable development.