---
title: "Role of institution and governance in the development"
author: "Eloïse Vannier"
format: pdf
execute:
  echo: false
---

```{r}
#| message: false
here::i_am("final_project.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(world)
library(spData)
library(raster)
```

```{r}
#| message: false
# Data loading :
govc_indicators <- vroom(here("data", "governance_indicators.csv"),na = (".."))
dvpt_indicators <- vroom(here("data", "development_indicators.csv"),na = (".."))

# Keeping only rows containing countries for which both data sets provide data :
govc_indicators = govc_indicators |> filter(nchar(`Country Code`)==3)
list_country = govc_indicators |> distinct (`Country Name`,`Country Code`)
dvpt_indicators = inner_join(list_country,dvpt_indicators,join_by(`Country Name`,`Country Code`))

# Changing the type of the values of the development indicators (perceived as characters before) :
dvpt_indicators <- dvpt_indicators |>  mutate(`2022 [YR2022]` = parse_double(`2022 [YR2022]`), `2021 [YR2021]` = parse_double(`2021 [YR2021]`),`2020 [YR2020]` = parse_double(`2020 [YR2020]`))

# Computing the mean of the three years to get one and only value for each indicators and have less NA :
mean.govc <- rowMeans((govc_indicators |> select(`2020 [YR2020]`,`2021 [YR2021]`, `2022 [YR2022]`)),na.rm = TRUE)
govc_indicators <- govc_indicators |> mutate("Mean"=mean.govc)
mean.dvpt <- rowMeans((dvpt_indicators |> select(`2020 [YR2020]`,`2021 [YR2021]`, `2022 [YR2022]`)),na.rm = TRUE)
dvpt_indicators <- dvpt_indicators |> mutate("Mean"=mean.dvpt)

# Saving the series' code in a separate dataframe and suppressing the "Series Code" column as well as the three "years column" in each dataframe :
govc_codes = govc_indicators |> distinct(`Series Name`,`Series Code`) 
dvpt_codes = dvpt_indicators |> distinct(`Series Name`,`Series Code`)

govc_indicators = govc_indicators |> select(-`Series Code`,-`2020 [YR2020]`,-`2021 [YR2021]`, -`2022 [YR2022]`)
dvpt_indicators = dvpt_indicators |> select(-`Series Code`,-`2020 [YR2020]`,-`2021 [YR2021]`, -`2022 [YR2022]`)

# Supressing the indicators providing data for less than 75% of the countries in our sample :
critic_minimum = 0.75*NROW(list_country)
keep.dvpt.indicator = " "
for (i in 1:NROW(dvpt_codes)) {
  df = dvpt_indicators |> filter(dvpt_indicators$`Series Name`== dvpt_codes$`Series Name`[i])
  if (sum(is.na(df$Mean)) < critic_minimum) {
    keep.dvpt.indicator = c(keep.dvpt.indicator,dvpt_codes$`Series Name`[i])
  }
}
keep.dvpt.indicator = as.data.frame(keep.dvpt.indicator) |>
  rename(`Series Name`=keep.dvpt.indicator)
dvpt_indicators <- inner_join(keep.dvpt.indicator,dvpt_indicators,join_by(`Series Name`))

keep.govc.indicator = " "
for (i in 1:NROW(govc_codes)) {
  df = govc_indicators |> filter(govc_indicators$`Series Name`== govc_codes$`Series Name`[i])
  if (sum(is.na(df$Mean)) < critic_minimum) {
    keep.govc.indicator = c(keep.govc.indicator,govc_codes$`Series Name`[i])
  }
}
keep.govc.indicator = as.data.frame(keep.govc.indicator) |>
  rename(`Series Name`=keep.govc.indicator)
govc_indicators <- inner_join(keep.govc.indicator,govc_indicators,join_by(`Series Name`))

# Joining the two dataframes :
wide_govc <- pivot_wider(govc_indicators, names_from = `Series Name`, values_from = Mean)
wide_dvpt <- pivot_wider(dvpt_indicators, names_from = `Series Name`, values_from = Mean)

global_df <- inner_join(wide_dvpt,wide_govc, join_by(`Country Name`,`Country Code`))

long_global_df <- pivot_longer(global_df, cols = `Access to electricity (% of population)`:`Voice and Accountability: Standard Error`, names_to = "Series Name", values_to = "Mean")
```

## I) Research question presentation

## II) Data sets descriptions

## III) Data analysis
When we think about "development", we first think of economic growth. If we want to improve the situation of developing countries, we must then try to foster growth. But how can we go that ? What parameters play a role in economic growth ?
While trying to explain economic growth, Sollow acknowledged two factors of production : capital and labor. The "residual" he obtained once he took those factors into account was significant and attributed to the productivity of the production factors. Several variables of the data set can indicate productivity : the HCI index computed by the World Bank for females, males and the total population, and the number of researchers in R&D per million people. As mentioned earlier the HCI was computed precisely to estimates the productivity of workers of tomorrow, and the more researchers there is, the more likely they will find new innovation that will possibly increase capital and/or labor productivity.
Let's check if our data go along with those hints :

```{r}
ggplot(global_df, aes(x =`GDP per capita growth (annual %)`, y = `Human capital index (HCI) (scale 0-1)`))+geom_point()+geom_smooth(method = "lm", se = FALSE, col = "red")
```
```{r}
pivot_longer(global_df,cols = c("Human capital index (HCI) (scale 0-1)","Human capital index (HCI), female (scale 0-1)","Human capital index (HCI), male (scale 0-1)"), 
                        names_to = "HCI", 
                        values_to = "HCI value") |>
  ggplot(aes(x =`GDP per capita, PPP (current international $)`,y = `HCI value`, group = HCI)) + 
  geom_point(aes(color = HCI))
```
These two first graphs show a slightly positive correlation between the HCI and the GDP growth, and a positive but decreasing relationship between the HCI and the GDP per capita taking into account the parity of purchasing power. We can also note that the HCI is higher for females than for males : that means that there is more chance for a girl than for a boy born between 2020 and 2022 to become as productive as they would if they enjoyed full health and complete education when they will grow up. 

```{r}
ggplot(global_df, aes(x =`GDP per capita, PPP (current international $)`, y = `Researchers in R&D (per million people)`))+geom_point()+geom_smooth(method = "lm", se = FALSE, col = "red")
ggplot(global_df, aes(x =`GDP per capita growth (annual %)`, y = `Researchers in R&D (per million people)`))+geom_point()+geom_smooth(method = "lm", se = FALSE, col = "red")
```
The two graphs right above show a positive and quite linear correlation between the number of researchers per million and the GDP per capita taking into account the parity of purchasing power. The relationship between the number of researchers and the economic growth is far less clear.
Of course those four graphs are not sufficient to deduce a causal relationship between those variables, especially to see the links between HCI or the number of researchers per million people and economic growth. Indeed since those variable are supposed to impact future growth we would need either the past number of researchers per million people and the HCI level of at least 20 years ago, or the future economic growth to see the eventual impacts. But since the corresponding data are not available, those results are still interesting : they at least do not seem not contradict our hypotheses.

## IV) Conclusion

## V) Annex (data import work and sources)

#### Data import work :
Here are the few steps to compute the global data frame used in this study :
First both data sets are loaded, we make sure that ".." is perceived as NA.
We check that the type of each of the column with values is double, and make the change necessary for the values'column of the development data set for which it was not the case.
Then the rows containing data of countries present in both data frame are selected.
In order to have a variable column with less NA, we compute the mean of those three columns.
We then select the indicators for which there is more than 25% of NA. 
Finally, the joining of the two data sets : we pivot them to get their wide version, and do an inner join.
Once they are joined, we pivot again the global data frame we just obtained to have its longer form.

The code of this preparation work is visible at the beginning of the quarto file corresponding to this report. A longer and more detailed version of it is also available in the "project_goal&data_description" quarto file present in the github repository of this project : [github-repository](https://github.com/E10is3/final-project.git)

#### Sources :
The data I use come from two open source data sets of the World Bank. 
The first one includes governance indicators available via this link : [WorldBank-governance-indicators](https://databank.worldbank.org/source/worldwide-governance-indicators#)
Please note that the ticked options are the preset ones by the website and not the ones I will use. To get the exact same .csv file, select all the 214 countries available, the years 2020, 2021 and 2022, and all the 36 variable series before the download.
The second data set includes development indicators, available via this link : [WorldBank-development-indicators](https://databank.worldbank.org/source/world-development-indicators#)
To get the exact same .csv file, select all the 266 countries available, the years 2020, 2021 and 2022. I first selected `r NROW(dvpt_codes)` variable series but kept only `r NROW(keep.dvpt.indicator)-1` of them, since the other three provided data for less than 75% of the countries of the final sample. Here is the list of those `r NROW(keep.dvpt.indicator)-1` variables series :
```{r}
#| message: false
print(keep.dvpt.indicator)
```













